### 2018-10-04 paper/finitecomplete-failures.tex
Fixed errors found by Wen-ling

For a given reference process $P$, its normalised transition graph
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
and each integer $p\ge 0$, we define a test for failures refinement as shown
below. 
%
\begin{eqnarray}
\label{eq:UFP}
U_F(p) & = & U_F(p,0,\ii n)
\\
U_F(p,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (k < p) \& \big(e:[n]^0   \then U_F(p,k+1,t(n,e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (k = p \wedge \minhits(n)\neq\varnothing) \&  \nonumber
\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}

% --------------------------------------------------------------------------
\subsubsection*{Explanation of the Test Definition}
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a \emph{test execution} is a trace of the
concurrent process
%\[
$Q\parallel[\Sigma] U_F(p)$.
%\]

It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
or $\epass$ always become visible, if they are the only events $U_F(p)/s$ is
ready to engage in: if $U_F(p)/s$ can only produce $\epass$ or $\efail$, the
occurrence of these events can never be blocked due to a livelock in $Q$
occurring in the same step of the execution.

The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and
only if, {\it every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with
the event $\epass$. This can also be  expressed by means of a failures
refinement as defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test} in the
literature, because every test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}.

We note that it is necessary to use failures refinement in the definition
above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(p)) \hide
\Sigma$ may have the same visible traces $\varepsilon$ and $\epass$ as the
``Test Passed Process'' $(\epass\then\Stop)$. However, the former may
nondeterministically refuse $\epass$, due to a deadlock occurring when a
faulty SUT process executes concurrently with $U_F(p,k,n)$ executing
branch~(\ref{eq:ufd}), when the guard condition $(k = p\wedge
\minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
in the next paragraphs.

Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
a length $p$. If, after having already run through $s\in\trc(P)$ with $\#s
\le p$, an event is accepted by the SUT that is outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice. %in the process
%$U_F(p,\#s,n)$ defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, this is
revealed by $\minhits(G(P)/s) = \varnothing$ (recall
(\ref{eq:minhitminaccempty}) and Lemma~\ref{lemma:ufproperties}). In this
case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
external choice in $U_F(p,\#s,G(P)/s)$). If $P/s$ may also
nondeterministically engage into events in such a situation, branch
(\ref{eq:ufc}) of the test is simultaneously enabled. If $Q/s$ is able to
engage into an event from $\Sigma - [P/s]^0$, a test execution exists where
$U_F(p,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
event.

If the length of $s$ is still less than $p$, the test accepts any event $e$
from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
$U_F(p,\#s+1,G(P)/s.e)$ in branch~(\ref{eq:ufc}); this follows again from
Lemma~\ref{lemma:ufproperties}. A test of this type is called
\emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts
its consecutive behaviour to the event selected by the SUT.

Finally, we suppose that a test execution has run through a trace
$s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n =
G(P)/s$. If $\minhits(n)\neq \varnothing$, the test changes its
behaviour:~instead of offering {\it all} legal events from $[n]^0$ to the
SUT, it nondeterministically chooses a minimal hitting set $H\in \minhits(n)$
and only offers the events contained in $H$. If the SUT refuses to engage
into some event of $H$, this reveals a violation of failures
refinement:~according to Lemma~\ref{lemma:hseta}, a conforming SUT should
accept at least one event of each minimal hitting set in $\minhits(n)$.
Therefore, the test execution only terminates with  $\epass$, if such an
event is accepted by the SUT. Otherwise, it deadlocks.

The specification of $U_F(p,k,n)$ implies that the test always stops after
having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$. If
branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
$p+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
is the last to be entered, the maximal length of $s$ is $p$, and the
execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
entered, the process either accepts another event $e$ of some minimal hitting
set $H\in\minhits(n)$ with $n = G(P)/s$ according to
Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $p+1$, and
the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
and the execution stops without a PASS or FAIL event. Such an execution is
also interpreted as FAIL, because it reveals that $(\epass\then\Stop)
\not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and $\minhits(n)$
are finite. Moreover, we further recall that $\minhits(n)$ may be empty, in
which case the indexed internal choice in (\ref{eq:ufd}) would be undefined.
The guard in that branch, however, requires $\minhits(n)\neq\varnothing$, and
branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.

%Before explaining the intuition behind this test definition in the paragraphs
%below, we state its basic properties in the following lemma, to show the
%relationships between $U_F(p)$ and the reference process $P$ from which the
%former has been derived. \fixme{For a reader, it is better to see the
%explanation first. Since this is easier to undo, I've changed it. }

We state basic properties of our tests in the following lemma, to show the
relationships between $U_F(p)$ and the reference process $P$ from which it is
derived.

\begin{lemma}\label{lemma:ufproperties}
If $s\in\trc(P)$ satisfies $\#s\le p$, then
$s, s.e\in\trc(U_F(p))$ for all $e\in\Sigma$, and the following properties hold.
\begin{eqnarray}
\label{eq:ifpa}
  &  & U_F(p)/s = U_F(p,\#s,G(P)/s)
\\
\label{eq:ifpb}
e\not\in [P/s]^0 & \implies & U_F(p)/s.e = (\efail\then\Stop)
\\
\label{eq:ifpc}
U_F(p)/s = U(p,\# s,n)  & \implies & [n]^0 = [P/s]^0
\\
\label{eq:ifpd}
U_F(p)/s = U(p,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
\end{eqnarray}
\end{lemma}
\begin{proof}
We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of
$G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
\le k < p$, so that $U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace
of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$ can
engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
defined, and $t(G(P)/s,e) = G(P)/s.e$. This leads to a new recursion, so that
$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ as required.

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(p)/s
= U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
$U_F(p,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
specification of this branch implies that
$$
U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = (\efail\then\Stop).
$$
%
Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$. Finally, statement
(\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because, according to
(\ref{eq:minhitsGP}), $\minhits(G(P)/s) = \minhits(P/s)$ for all traces of
$P$. \xbox
\end{proof}
%
Note that it is not guaranteed for $U_F(p)$ to run through the traces $s,
s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$
event by entering branch (\ref{eq:ufb}). Therefore,
Lemma~\ref{lemma:ufproperties} just states the existence of
$U_F(p)$-executions $s, s.e$ satisfying the properties stated there.

@@

A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
\lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
over $P$'s alphabet that may or may not conform to $P$.

@@

% -------------------------------------------------------------------------
\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{$U_F(j)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
% ------------------------------------------------------------------------
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
A test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}We make two points in separate steps below. The
first is that the test execution cannot reach branch~(\ref{eq:ufa}) and raise
a $fail$ event.  The second is that it cannot deadlock without raising a
$pass$ event. This case would also be interpreted as FAIL, since then
$\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
U_F(k))\hide \Sigma$.
% alcc: part of the explanation above was in the following lemma, but it is already
% relevant here.

@@

\paragraph{Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
that $U_F(j)/s = U_F(j,k,n)$ with $U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so
$\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
$U_F(j,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
\varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
without a verdict. \fixme{I think we need a stronger argument here, as to why
it cannot deadlock.}
\fxnote{jp: I do not understand the question -- let's discuss this on the phone.}
 For the last branch, we note that it can only be entered
if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
minimal hitting set $H\in\minhits(n)$ has a non-empty intersection with each
of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ never blocks
when offered events from $H\in\minhits(n)$, and the test terminates with
event $\epass$. Note that this argument requires that $Q$ is free of
livelocks, because otherwise the $\epass$-events might not become visible,
due to unbounded sequences of hidden events performed by $Q$. \xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there and for conformance relation $\lessdet_F$.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the $\efail$
event or deadlocks. We do this by analysing the product graph of the reference
process $P$ and the SUT process $Q$: any trace $s\in\trc(Q)\cap\trc(P)$ gives rise to 
a path labelled by the events of $s$ 
through this product graph. Any error can be detected after running through such a trace
and then either observing an event outside $[P/s]^0$ (this is the violation described
by Case~1) or identifying an illegal acceptance $A_Q$ as specified in Case~2. 
It is not guaranteed, however, that trace $s$ is short enough to be executed by one of the 
test cases $U_F(j)$ with $0\le j < pq$. Therefore, it has to be shown that for any $s$ leading to an error
situation specified by Case~1 or Case~2, there exists a trace $u$ of maximal length $pq-1$
leading to the same error.

>>>>

For a given reference process $P$, its normalised transition graph
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
and each integer $j\ge 0$, we define a test for failures refinement as shown
below. 
%
\begin{eqnarray}
\label{eq:UFP}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (k < j) \& \big(e:[n]^0   \then U_F(j,k+1,t(n,e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \&  \nonumber
\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}

% --------------------------------------------------------------------------
\subsubsection*{Explanation of the Test Definition}
A test is performed by running $U_F(j)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a \emph{test execution} is a trace of the
concurrent process
%\[
$Q\parallel[\Sigma] U_F(j)$.
%\]

It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
or $\epass$ always become visible, if they are the only events $U_F(j)/s$ is
ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
occurrence of these events can never be blocked due to a livelock in $Q$
occurring in the same step of the execution.

The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(j)$) if, and
only if, {\it every} execution of $Q\parallel[\Sigma] U_F(j)$ terminates with
the event $\epass$. This can also be  expressed by means of a failures
refinement as defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(j) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test} in the
literature, because every test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}.

We note that it is necessary to use failures refinement in the definition
above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide
\Sigma$ may have the same visible traces $\varepsilon$ and $\epass$ as the
``Test Passed Process'' $(\epass\then\Stop)$. However, the former may
nondeterministically refuse $\epass$, due to a deadlock occurring when a
faulty SUT process executes concurrently with $U_F(j,k,n)$ executing
branch~(\ref{eq:ufd}), when the guard condition $(k = j\wedge
\minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
in the next paragraphs.
Alternatively, a faulty SUT $Q$ 
might internally deadlock after a trace $s$ where $\#s < j$ and
$\minhits(G(P)/s) \neq \varnothing$,
so that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.

Intuitively speaking, $U_F(j)$ is able to perform any trace $s$ of $P$, up to
a length $j$. If, after having already run through $s\in\trc(P)$ with $\#s
\le j$, an event is accepted by the SUT that is outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(j)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice. %in the process
%$U_F(j,\#s,n)$ defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, this is
revealed by $\minhits(G(P)/s) = \varnothing$ (recall
(\ref{eq:minhitminaccempty}) and Lemma~\ref{lemma:ufproperties}). In this
case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
external choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also
nondeterministically engage into events in such a situation, branch
(\ref{eq:ufc}) of the test is simultaneously enabled. If $Q/s$ is able to
engage into an event from $\Sigma - [P/s]^0$, a test execution exists where
$U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
event.

If the length of $s$ is still less than $j$, the test accepts any event $e$
from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
$U_F(j,\#s+1,G(P)/s.e)$ (note that $G(P)/s.e = t(G(P)/s,e)$) 
in branch~(\ref{eq:ufc}); this follows again from
Lemma~\ref{lemma:ufproperties}. A test of this type is called
\emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts
its consecutive behaviour to the event selected by the SUT.

Now suppose that a test execution has run through a trace
$s\in\trc(P)$ of length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n =
G(P)/s$. If $\minhits(n)\neq \varnothing$, the test changes its
behaviour:~instead of offering {\it all} legal events from $[n]^0$ to the
SUT, it nondeterministically chooses a minimal hitting set $H\in \minhits(n)$
and only offers the events contained in $H$. If the SUT refuses to engage
into some event of $H$, this reveals a violation of failures
refinement:~according to Lemma~\ref{lemma:hseta}, a conforming SUT should
accept at least one event of each minimal hitting set in $\minhits(n)$.
Therefore, the test execution only terminates with  $\epass$, if such an
event is accepted by the SUT. Otherwise, it deadlocks, and the test fails.

The specification of $U_F(j,k,n)$ implies that the test always stops after
having engaged into a trace $s\in\trc(Q)$ of maximal length $j$ or $j+1$. If
branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
$j+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
is the last to be entered, the maximal length of $s$ is $j$, and the
execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
entered, the process either accepts another event $e$ of some minimal hitting
set $H\in\minhits(n)$ with $n = G(P)/s$ according to
Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $j+1$, and
the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,  the final length of $s$ is $j$,
and the execution stops without a PASS or FAIL event. Such an execution is
also interpreted as FAIL, because it reveals that $(\epass\then\Stop)
\not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(j)$ is finite, because the number of traces $s$ with maximal length
$(j+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and $\minhits(n)$
are finite. Moreover, we further recall that $\minhits(n)$ may be empty, in
which case the indexed internal choice in (\ref{eq:ufd}) would be undefined.
The guard in that branch, however, requires $\minhits(n)\neq\varnothing$, and
branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.

We state basic properties of our tests in the following lemma, to show the
relationships between $U_F(j)$ and the reference process $P$ from which it is
derived.

\begin{lemma}\label{lemma:ufproperties}
If $s\in\trc(P)$ satisfies $\#s\le j$, then
$s, s.e\in\trc(U_F(j))$ for all $e\in\Sigma$, and the following properties hold.
\begin{eqnarray}
\label{eq:ifpa}
  &  & U_F(j)/s = U_F(j,\#s,G(P)/s)
\\
\label{eq:ifpb}
e\not\in [P/s]^0 & \implies & U_F(j)/s.e = (\efail\then\Stop)
\\
\label{eq:ifpc}
U_F(j)/s = U_F(j,\# s,n)  & \implies & [n]^0 = [P/s]^0
\\
\label{eq:ifpd}
U_F(j)/s = U_F(j,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
\end{eqnarray}
\end{lemma}
\begin{proof}
We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
the statement holds because $U_F(j)$ starts with the initial node $\ii n$ of
$G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
\le k < j$, so that $U_F(j)/s = U_F(j,\#s,G(P)/s)$. Now let $s.e$ be a trace
of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(j,\#s,G(P)/s)$ can
engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
defined, and $t(G(P)/s,e) = G(P)/s.e$. This leads to a new recursion, so that
$U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = U_F(j,\#s+1,G(P)/s.e)$ as required.

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
specification of this branch implies that
$$
U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = (\efail\then\Stop).
$$
%
Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$. Finally, statement
(\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because, according to
(\ref{eq:minhitsGP}), $\minhits(G(P)/s) = \minhits(P/s)$ for all traces of
$P$. \xbox
\end{proof}
%
Note that it is not guaranteed for $U_F(j)$ to run through the traces $s,
s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
some prefix $u$ of $s$: in such a case, $U_F(j)$ may stop with a $\epass$
event by entering branch (\ref{eq:ufb}). Therefore,
Lemma~\ref{lemma:ufproperties} just states the existence of
$U_F(j)$-executions $s, s.e$ satisfying the properties stated there.

@@

A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ counsists of a
reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
\lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
over $P$'s alphabet that may or may not conform to $P$.

@@

% -------------------------------------------------------------------------
\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{$U_F(j)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
% ------------------------------------------------------------------------
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
A test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}We make two points in separate steps below. The
first is that the test execution cannot reach branch~(\ref{eq:ufa}) and raise
a $fail$ event.  The second is that it cannot deadlock without raising a
$pass$ event. This case would also be interpreted as FAIL, since then
$\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
U_F(j))\hide \Sigma$.
% alcc: part of the explanation above was in the following lemma, but it is already
% relevant here.

@@

\paragraph{Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
that $U_F(j)/s = U_F(j,k,n)$ with $U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so
$\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
$U_F(j,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
\varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
without a verdict. \fixme{I think we need a stronger argument here, as to why
it cannot deadlock.}
\fxnote{jp: I do not understand the question -- let's discuss this on the phone.}
 For the last branch, we note that it can only be entered
if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
minimal hitting set $H\in\minhits(n)$ has a non-empty intersection with each
of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ never blocks
when offered events from $H\in\minhits(n)$, and the test terminates with
event $\epass$. Note that this argument requires that $Q$ is free of
livelocks, because otherwise the $\epass$-events might not become visible,
due to unbounded sequences of hidden events performed by $Q$. \xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there and for conformance relation $\lessdet_F$.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
event or deadlocks. We do this by analysing the product graph of the reference
process $P$ and the SUT process $Q$: any trace $s\in\trc(Q)\cap\trc(P)$ gives rise to 
a path labelled by the events of $s$ 
through this product graph. Any error can be detected after running through such a trace
and then either observing an event outside $[P/s]^0$ (this is the violation described
by Case~1) or identifying an illegal acceptance $A_Q$ as specified in Case~2. 
It is not guaranteed, however, that trace $s$ is short enough to be executed by one of the 
test cases $U_F(j)$ with $0\le j < pq$. Therefore, it has to be shown that for any $s$ leading to an error
situation specified by Case~1 or Case~2, there exists a trace $u$ of maximal length $pq-1$
leading to the same error.
